Retrofitting Effect Handlers onto OCaml

### 2020-10-29 retro-concurrency.tex
Updates

The |Read| effect can then be handled by another handler:

@@

\begin{enumerate}[label=R\arabic*]
	\item \textbf{Backwards compatibility} A well behaved OCaml program does not
		break under OCaml extended with effect handlers. OCaml code that does not
		use effect handlers will pay negligible performance cost.
	\item \textbf{Tool compatibility} OCaml with effect handlers produce
		well-formed backtraces, and remains compatible with program analysis tools
		such as debuggers and profilers that inspect the stack using DWARF unwind
		tables.
	\item \textbf{Effect handler efficiency} In order to support
		highly-concurrent applications, the program must accommodate millions of
		continuations in the heap. Installing effect handlers, capturing and
		resuming continuations must be fast.
	\item \textbf{Effect handler expressivity} While making continuations
		resumable at-most once (\emph{affine}) is a standard way to make
		continuation capture fast, we must still support use cases where the
		continuations are resumed more than once.
	\item \textbf{Forwards compatibility} While this work does not present an
		effect system, the design must be compatible with a future effect system
		for OCaml. As a cornerstone of modularity, we also want blocking I/O code
		to \emph{transparently} be made asynchronous with the help of effect
		handlers.
\end{enumerate}

@@

\begin{itemize}
	\item the design and implementation of affine effect handlers for OCaml with
		negligible performance impact on code that does not use effect handlers.
		Our design lets us retain OCaml's compatibility with program analysis tools
		that inspect the stack.
	\item a formal operation semantics for the effect handler implementation in
		OCaml.
	\item extensive evaluation which shows that our implementation has negligible
		impact on code that does not use effect handlers, and serves as an
		efficient foundation for scalable concurrent and parallel programming.
\end{itemize}

@@

\subsection{Discontinuing a continuation}

The interaction of non-local control flow with systems programming is quite
subtle~\cite{}. Consider the following function that uses blocking I/O
functions from the OCaml standard library to copy data from the input channel
|ic| to the output channel |oc|:

@@

\subsection{Multi-shot continuations}

@@

The function |bar| expects |foo| to return exactly once, ignoring any exceptions. However,
due to continuation cloning, the function returns twice, leading to double-free
error.

>>>>

The |Read| effect can then be handled by another handler enclosing the
division-by-zero handler:

@@

\begin{enumerate}[label=R\arabic*]
	\item \textbf{Backwards compatibility} A well behaved OCaml program does not
		break under OCaml extended with effect handlers. OCaml code that does not
		use effect handlers will pay negligible performance cost.
	\item \textbf{Tool compatibility} OCaml with effect handlers produce
		well-formed backtraces, and remains compatible with program analysis tools
		such as debuggers and profilers that inspect the stack using DWARF unwind
		tables.
	\item \textbf{Effect handler efficiency} In order to support
		highly-concurrent applications, the program must accommodate millions of
		continuations in the heap. Installing effect handlers, capturing and
		resuming continuations must be fast.
\if{0}
	\item \textbf{Effect handler expressivity} While making continuations
		resumable at-most once (\emph{affine}) is a standard way to make
		continuation capture fast, we must still support use cases where the
		continuations are resumed more than once.
\fi
	\item \textbf{Forwards compatibility} While this work does not present an
		effect system, the design must be compatible with a future effect system
		for OCaml. As a cornerstone of modularity, we also want blocking I/O code
		to \emph{transparently} be made asynchronous with the help of effect
		handlers.
\end{enumerate}

@@

\begin{itemize}
	\item the design and implementation of affine effect handlers for OCaml. Our
		design retains OCaml's compatibility with program analysis tools that
		inspect the stack using DWARF unwind tables. We have validated our DWARF
		unwind tables using an automated validator tool~\cite{}.
	\item a formal operation semantics for the effect handler implementation in
		OCaml.
	\item extensive evaluation which shows that our implementation has negligible
		impact on code that does not use effect handlers, and serves as an
		efficient foundation for scalable concurrent and parallel programming.
\end{itemize}

@@

The interaction of non-local control flow with systems programming is quite
subtle~\cite{TFP17}. Consider the following function that uses blocking I/O
functions from the OCaml standard library to copy data from the input channel
|ic| to the output channel |oc|:

@@

\if{0}
\subsection{Multi-shot continuations}

@@

The function |bar| expects |foo| to return exactly once, ignoring any exceptions. However,
due to continuation cloning, the function returns twice, leading to double-free
error.
\fi
